Review and comparison of tools for managing X.509 type keys and certificates

### 2016-12-01 TEX/draft.tex
Only Keystore Explorer is left

\section{General}
\textit{Tested under Fedora 21, OpenSSL version 1.0.1k-fips} %DOPLNIT INFO O CNF FILE
  \subsection{Type}
  OpenSSL uses PKCS12 to store keys and certificates. Certificates and keys generated in OpenSSL are being made into PEM or DER encoded file. PEM or DER encoded keys and certificates can then be stored inside single PKCS\#12 file.

@@

  \subsection{License}
OpenSSL is free to use commercialy, however creating CA is not advised in O'REILLY's Network Security with OpenSSL by J. Viega, M. Messier and P. Chandra on page 59,
\begin{displayquote}
\textit{Since OpenSSL's command-line CA functionality was intended primarely as an example of how to use OpenSSL to build a CA, we don't recommend that you attempt to use it in a large production environment.}
\end{displayquote}
and also hinted in manual pages:
\begin{displayquote}
\textit{The ca command is quirky and at times downright unfriendly.}

@@

    \subsubsection{Create combination of private key and signed chain}
    First creation of the request is needed, after that request must be signed. To combine them together the creation of PKCS\#12 file is needed with commands:
\begin{verbatim}
	openssl pkcs12 -export -out outfilename.p12 \
	        -in signed_certificate.crt -inkey privateKey.key \
	        -chain -CAfile ca-all.crt -password pass:PASSWORD
\end{verbatim}
    \subsubsection{Specify certificate validity}
    Time how long the certificate will be valid can be given in csr but only as a suggestion. Ultimately the real time of validity is always given by CA signing the CSR.
    \subsubsection{Setting Subject Alternative Name for end certificates}
    The Subject Alternative Name is also given in \verb+[ v3_req ]+ part of openssl.cnf that CA uses. It can be given as an IP address but also as a DNS name.

@@

    \subsubsection{Certificate Signing Request signing}
Possible %but i need to look wy it does not work for me
    \subsubsection{Create combination of private key and signed chain}
When you create a certificate by signing a request to sign key only, you can validate chain. %but i think that is not what is regarded here as signed chain    
    \subsubsection{Specify certificate validity}
Validity is specified by changing cert.cfg file used while generating a certificate, the part \verb+expiration_days = <value>+
Alternatively, you cen set \verb+NOT BEFORE+ and \verb+NOT AFTER+ times by changing \verb+activation_date = <NOT_BEFORE_STRING>+ and \verb+expiration_date = <NOT_AFTER_STRING>+ where the times are written in GNU date string format.

@@

    \subsubsection{Certificate or certificate chain from a file}

    \subsubsection{Private key only}
It is not possible to export private key only, however for proper use of GnuTLS keys need to be stored on drive or imported to PKCS\#12 file.
  \subsection{Importing certificates and keys into storage files}
Possible.

@@

\section{OpenSSL}
It is a great command line tool, can work with PEM, DER and PKCS\#12 files. Strong point is an easy form of creating a self-signed certificate, however using it for CA certificates with ca option is not ideal - it is noted in manual pages and also a O'REILLY OpenSSL book.
Another strong point is configuration, which can be done in file. I tested it with creation of appropriate directories and configs.

>>>>

\section{General}
\textit{Tested under Fedora 21, OpenSSL version 1.0.1k-fips} %DOPLNIT INFO O CNF FILE

OpenSSL heavily relies on a configuration file that needs to be prepared along with a default directory sequence in advance. Hierarchy of that directory sequence can be seen in supplied OpenSSL.cnf file under \verb+[ CA_default ]+ part.

Time how long the certificate will be valid can be given in csr but only as a suggestion. Ultimately the real time of
validity is always given by CA signing the CSR.

  \subsection{Type}
  OpenSSL uses PKCS12 to store keys and certificates. Certificates and keys generated in OpenSSL are being made into PEM or DER encoded file. PEM or DER encoded keys and certificates can then be stored inside single PKCS\#12 file.

@@

  \subsection{License}
OpenSSL is free to use commercialy. Full text is available at \url{https://www.openssl.org/source/license.html}  
\section{Generation and signing of certificates}

@@

    \subsubsection{Create combination of private key and signed chain}
    First creation of the request is needed, after that request must be signed. To combine them together the creation of PKCS\#12 file is needed with commands:
\begin{verbatim}
	openssl pkcs12 -export -out outfilename.p12 \
	        -in signed_certificate.crt -inkey privateKey.key \
	        -chain -CAfile ca-all.crt -password pass:PASSWORD
\end{verbatim}
    \subsubsection{Specify certificate validity}
    Time how long the certificate will be valid is chosen by CA during signing the CSR, where in command for signing certificate signing request we choose desired duration in days by changing \verb+-days <velidity_time>+ part.

    \subsubsection{Setting Subject Alternative Name for end certificates}
    The Subject Alternative Name is also given in \verb+[ v3_req ]+ part of openssl.cnf that CA uses. It can be given as an IP address but also as a DNS name.

@@

    \subsubsection{Certificate Signing Request signing}
Possible by command:
\begin{verbatim}
certtool --generate-certificate --load-request <csr_file> \
         --load-ca-certificate <path_to_ca_cert> \
         --load-ca-privkey <ca_cert_privkey_file> 
         --outfile <name>.pem
\end{verbatim}

    \subsubsection{Specify certificate validity}
Validity is specified by changing cert.cfg file used while generating a certificate, the part \verb+expiration_days = <value>+
Alternatively, you cen set \verb+NOT BEFORE+ and \verb+NOT AFTER+ times by changing \verb+activation_date = <NOT_BEFORE_STRING>+ and \verb+expiration_date = <NOT_AFTER_STRING>+ where the times are written in GNU date string format.

@@

    \subsubsection{Private key only}
It is not possible to export private key only, however for proper use of GnuTLS keys need to be stored on drive or imported to PKCS\#12 file.
  \subsection{Importing certificates and keys into storage files}
To import certificate and key into a PKCS\#12 file use command:
\begin{verbatim}
certtool --load-certificate <in_cert> \
         --load-privkey <in_key> \
         --to-p12 --outder --outfile <name>.[p12|pfx]
\end{verbatim}

@@

\section{OpenSSL}
It is a great command line tool, can work with PEM, DER and PKCS\#12 files. Strong point is an easy form of creating a self-signed certificate, however using it for CA certificates with ca option is not ideal -  creating CA is not advised in O'REILLY's Network Security with OpenSSL by J. Viega, M. Messier and P. Chandra on page 59,
\begin{displayquote}
\textit{Since OpenSSL's command-line CA functionality was intended primarely as an example of how to use OpenSSL to build a CA, we don't recommend that you attempt to use it in a large production environment.}
\end{displayquote}
and also hinted in manual pages:
\begin{displayquote}
\textit{The ca command is quirky and at times downright unfriendly.}

On the other hand strong point of OpenSSL is configuration of created certificate, which can be done in file. I tested it with creation of appropriate directories and configs.
